On the number of tiles visited by a\\ line segment on a rectangular grid

### 2021-11-27 paper.tex
Tried to clarify proof of Theorem 1
The explanation of the three cases for the solution pair (i+,j+) were confusing. I think it's clearer now, but more work is still needed

The pair $(\isolr, \jsolr)$ can be obtained as a solution of the equation system
\begin{align}
\label{eq: icont jcont 1}
(\isolr-2)^2 a^2 + (\jsolr-2)^2 b^2 &= \len^2, \\
\label{eq: icont jcont 2}
\jsolr = \frac{\isolr a^2}{b^2} - \frac{5a^2}{2b^2} + \frac 3 2.
\end{align}
Substituting \eqref{eq: icont jcont 2} into \eqref{eq: icont jcont 1} yields a quadratic equation for $\isolr$,
\begin{equation}
\label{eq: quadratic, i}
4a^2(\isolr-2)^2 - 4a^2(\isolr-2) + a^2+b^2-\frac{4\len^2 b^2}{a^2+b^2} = 0,
\end{equation}
which can have zero, one or two real-valued solutions. These cases respectively mean that the solid line in Figure~\ref{fig: ijLS_1p35_detail} is exterior, tangent or secant to the circle (the figure shows the latter case). Only the solution with $\isolr, \jsolr \geq 2$, if any, is of interest. This solution can only exist as the one with largest $\isolr$ value when there are real-valued solutions, and it is obtained as %***Here rewriting
\begin{equation}
\label{eq: icont jcont 3}
\isolr = \frac{a + b \sqrt{4\len^2/(a^2+b^2)-1}}{2a} + 2.
\end{equation}

The case with no real-valued solutions to \eqref{eq: quadratic, i} corresponds to $\len < \sqrt{a^2+b^2}/2$. Since $a \geq b$, this implies that $\len < a$. Thus any achievable $(i,j)$ pair, if any, will have $i=2$. Therefore in this case $\isoli$ should be set to $2$; and then the maximum achievable $\jsoli$ will be obtained from \eqref{eq: icont jcont 2}.

If \eqref{eq: quadratic, i} has two real-valued solutions or one real-valued double solution, which occurs when $\len \geq \sqrt{a^2+b^2}/2$, the largest solution will have $\jsolr < 2$ if $\isolr < (b^2/a^2+5)/2$, as can be seen setting $\jsolr=2$ in \eqref{eq: icont jcont 2}; or equivalently if $\sqrt{a^2+b^2}/2 \leq \len < (a^2+b^2) / (2a)$, as can be seen substituting $\jsolr < 2$ and $\isolr < (b^2/a^2+5)/2$ in \eqref{eq: icont jcont 1}. Since $\isolr < (b^2/a^2+5)/2 < 3$ for $a \geq b$, only pairs with $i=2$ are achievable in this case again, and thus the solution $\isoli$ should also be $2$.

Lastly, for $\len \geq (a^2+b^2) / (2a)$ the expressions \eqref{eq: icont jcont 2} and \eqref{eq: icont jcont 3} give $\isolr, \jsolr \geq 2$, and $\isoli$ should be taken as the greatest integer less than $\isolr$, i.e.~$\lceil \isolr \rceil-1$.

The three cases are unified by taking the real part of \eqref{eq: icont jcont 3} and computing $\isoli = \lceil \isolr \rceil-1$, as is easily checked. This gives \eqref{eq: theo: funt, form; i}. Once $\isoli$ is known, $\jsoli$ is computed, according to \eqref{eq: theo: funt, form; j}, as the greatest integer such that $(\isoli a,\jsoli b)$ is within the circle determined by $\len$. This ensures that $(\isoli,\jsoli)$ is achievable with lengths less than $\len$.

There are two possibilities for the pair $(\isoli,\jsoli)$, as stated at the outset. These are illustrated in Figures~\ref{fig: ijLS_1p35_detail_in} and \ref{fig: ijLS_1p35_detail_notin} respectively. The first possibility is that $(\isoli,\jsoli) \in \mss$ (Figure~\ref{fig: ijLS_1p35_detail_in}). Then, by construction  $(\isoli,\jsoli)$ maximizes $i+j-1$ among all achievable pairs of $\mss$, and is therefore optimal.

>>>>

The pair $(\isolr, \jsolr)$ can be obtained as a solution of the equation system
\begin{align}
\label{eq: icont jcont 1}
(\isolr-2)^2 a^2 + (\jsolr-2)^2 b^2 &= \len^2, \\
\label{eq: icont jcont 2}
\jsolr = \frac{\isolr a^2}{b^2} - \frac{5a^2}{2b^2} + \frac 3 2.
\end{align}
Substituting \eqref{eq: icont jcont 2} into \eqref{eq: icont jcont 1} gives a quadratic equation for $\isolr$,
\begin{equation}
\label{eq: quadratic, i}
4a^2(\isolr-2)^2 - 4a^2(\isolr-2) + a^2+b^2-\frac{4\len^2 b^2}{a^2+b^2} = 0,
\end{equation}
which is solved as
\begin{equation}
\label{eq: icont jcont 3}
\isolr = \frac{a \pm b \sqrt{4\len^2/(a^2+b^2)-1}}{2a} + 2,
\end{equation}
and then $\jsolr$ can be obtained from \eqref{eq: icont jcont 2}. This can produce zero, one or two real-valued solution pairs $(\isolr, \jsolr)$, which respectively corresponds to the solid line in Figure~\ref{fig: ijLS_1p35_detail} being exterior, tangent or secant to the circle (the figure shows the latter case).

A solution $(\isolr, \jsolr)$ is of interest only if it is real-valued with $\isolr, \jsolr \geq 2$. This solution, if it exists, is always associated with the positive sign in \eqref{eq: icont jcont 3} (the negative sign either yields the same solution, in the tangent case, or yields a solution with $\jsolr < 2$, in the secant case). In this regard, three cases need to be distinguished: there are no real-valued solution pairs $(\isolr, \jsolr)$; there are one or two but do not satisfy $\isolr, \jsolr \geq 2$; there are one or two and and one of them satisfies that inequality.

If $\len < \sqrt{a^2+b^2}/2$, \eqref{eq: icont jcont 3} yields no real-valued solutions for $\isolr$. Since $a \geq b$, $\len < \sqrt{a^2+b^2}/2$ implies that $\len < a$, and therefore any achievable $(i,j)$ pair, if any, will have $i=2$. Thus in this case $\isoli$ should be set to $2$; and then the maximum achievable $\jsoli$ will be obtained from \eqref{eq: icont jcont 2}.

For $\len \geq \sqrt{a^2+b^2}/2$ there are either two real-valued solutions or one real-valued double solution for $\isolr$. These solutions will give $\jsolr < 2$ if $\isolr < (b^2/a^2+5)/2$, as can be seen setting $\jsolr=2$ in \eqref{eq: icont jcont 2}; or equivalently if $\sqrt{a^2+b^2}/2 \leq \len < (a^2+b^2) / (2a)$, as can be seen substituting $\jsolr < 2$ and $\isolr < (b^2/a^2+5)/2$ in \eqref{eq: icont jcont 1}. Thus for $\len$ in this range, since $\isolr < (b^2/a^2+5)/2 < 3$ for $a \geq b$, only pairs with $i=2$ are achievable again, and therefore $\isoli$ must also be $2$.

Lastly, for $\len \geq (a^2+b^2) / (2a)$ the expressions \eqref{eq: icont jcont 2} and \eqref{eq: icont jcont 3} with positive sign give $\isolr, \jsolr \geq 2$, and $\isoli$ should be taken as the greatest integer less than $\isolr$, i.e.~$\lceil \isolr \rceil-1$.

The three cases are unified by taking the real part of \eqref{eq: icont jcont 3} and computing $\isoli = \lceil \isolr \rceil-1$, as can be easily checked. This yield \eqref{eq: theo: funt, form; i}. Once $\isoli$ is known, $\jsoli$ is computed, according to \eqref{eq: theo: funt, form; j}, as the greatest integer such that $(\isoli a,\jsoli b)$ is within the circle determined by $\len$. This ensures that $(\isoli,\jsoli)$ is achievable with lengths less than $\len$.

There are two possibilities for the resulting pair $(\isoli,\jsoli)$, as stated at the outset. These are illustrated in Figures~\ref{fig: ijLS_1p35_detail_in} and \ref{fig: ijLS_1p35_detail_notin} respectively. The first possibility is that $(\isoli,\jsoli) \in \mss$ (Figure~\ref{fig: ijLS_1p35_detail_in}). Then, by construction  $(\isoli,\jsoli)$ maximizes $i+j-1$ among all achievable pairs of $\mss$, and is therefore optimal.
